---
title: "STA 440 Lab - Spatial Data Visualization"
author: "Anna Yanchenko"
date: "3/19/2021"
output: pdf_document
---

Lab GitHub: https://github.com/sta440-sp21/lab-spatial-319





```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message=FALSE, warning = FALSE}
library(tidyverse)
library(sf)
```

```{r}
## Load NC county data
nc <- st_read("nc.shp", quiet = TRUE)
nc
```

```{r}
## Plot
ggplot(nc) + 
  geom_sf() + 
  labs(title = "North Carolina county boundaries") +
  theme_bw()
```

```{r}
## read in health data

nc_health <- read_csv("nc_health.csv")
head(nc_health)
```

```{r}
nc_health <- rename(nc_health, name = county)
nc <- merge(nc, nc_health, by = "name")
nc
```



# 1. Create a choropleth map for a variable of your choice

- What type of variable do we need for a choropleth map?

\newpage


```{r}
ggplot(nc) + 
  geom_sf(aes(fill = inactive)) +
  scale_fill_gradient(low = "#fee8c8", high = "#7f0000") +
  labs(title = "Percentage of Adults who are Physically Inactive",
       fill = "Percentage") +
  theme_bw()
```


```{r}
ggplot(nc) + 
  geom_sf(aes(fill = obesity)) +
  scale_fill_gradient(low = "#fee8c8", high = "#7f0000") +
  labs(title = "Percentage of Adults who are Obese",
       fill = "Percentage") +
  theme_bw()
```


```{r}
ggplot(nc) + 
  geom_sf(aes(fill = hs)) +
  scale_fill_gradient(low = "#fee8c8", high = "#7f0000") +
  labs(title = "Percentage of Adults who have Graduated High School",
       fill = "Percentage") +
  theme_bw()
```


# 2. Evaluate whether there is evidence of spatial autocorrelation for obesity

- What measure can we use to access spatial autocorrelation?
- What type of neighbors do we want to consider?
- How do we interpret the results?

\newpage

```{r,  message=FALSE, warning = FALSE}
library(spdep)
```

```{r}
## Form spatial weight matrix
sp_wts <- poly2nb(nc, row.names=nc$name, queen = T)
sp_wts
summary(sp_wts)

## Spatial weight matrix
sp_mat <- nb2mat(sp_wts, style='B') # Binary 1/0

## List form of spatial weight matrix
sp_mat_list <-  nb2listw(sp_wts, style='B')
sp_mat_list



## moran(x, listw, n, S0)

moran(nc$obesity, sp_mat_list, nrow(nc), sum(sp_mat))

```

```{r}
sp_mat_list_std <-  nb2listw(sp_wts, style='W')
moran.plot(nc$obesity, sp_mat_list_std,
           xlab = "Percentage Obese",
           ylab = "Spatially lagged percentage")
```

# 3. Create a new variable that corresponds to the average obesity in neighboring counties (i.e., the spatially-lagged obesity). In your spatial weight matrix, use a row-standardized binary matrix.



\newpage

```{r}
nc$lagged_obesity <- t(nc$obesity %*% nb2mat(sp_wts, style='W'))
```


```{r}
ggplot(nc) + 
  geom_sf(aes(fill = lagged_obesity)) +
  scale_fill_gradient(low = "#fee8c8", high = "#7f0000") +
  labs(title = "Averge Percentage Obese of Neighbors",
       fill = "Percentage") +
  theme_bw()
```


```{r}
ggplot(nc) + 
  geom_sf(aes(fill = obesity)) +
  scale_fill_gradient(low = "#fee8c8", high = "#7f0000") +
  labs(title = "Percentage of Adults who are Obese",
       fill = "Percentage") +
  theme_bw()
```


```{r, echo = FALSE}
## Moran's I
#moran(nc$lagged_obesity, sp_mat_list, nrow(nc), sum(sp_mat))

```

# 4. Evaluate the Pearson correlation between a county's adult physical inactivity rate and the average obesity among its neighbors

\newpage

```{r}
cor(nc$inactive, nc$lagged_obesity, method = 'pearson')[1,1]
```

# Challenge: Find county-level population data; modify your spatial weight matrix to account for the population of neighboring counties.


- How might we do this?

\newpage

https://demography.osbm.nc.gov/explore/dataset/county-population-estimates-standard-revised/table/?disjunctive.county&disjunctive.population&sort=county&refine.year=2019
